Does an earthquake “know” how big it will be? A neural-net aided study

### 2024-05-20 article.tex
mostly typos

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/combined_batplots.pdf}
    \caption{
        \textbf{Metric scores for the MAGNET model, and common benchmark models. a}, \textbf{b}, \textbf{c}, Minus mean information content of our MAGNET model (red) and other common benchmark magnitude predictors (see labels on bars in the figure). A lower score indicated a better preforming model. \textbf{d, e, f}, Reciever Operating Characteristic (ROC) and \textbf{g, h, i} the interpolated precision-recall (PR) curve for a binary classifier determining the existence of a large ($m>=4$) event. The preformance of such a classifier can be quantified by the area under the curve (AUC), noted in each frame, colorcoded identically to the bar plots. For the AUC metrics, a higher score indicated a better preforming classifier.
        }
        \label{fig:metrics}
\end{figure}

@@

\subsection{Measurement of temporal incompleteness}
Temporal incompleteness is measured at the times of events by fitting the completeness magnitude to a window of 300 events: 150 past events, 1 current event, 149 future events. The window is constructed by the same algorithm, described to construct the \textit{ Last $n$ events benchmark}.

\section*{Data Availability}
This is a required section. Guidelines for data availability: \url{https://www.nature.com/documents/nr-data-availability-statements-data-citations.pdf}.

\section*{Code Availability}
\textbf{Grey's Note} \textit{This is a required section. If your article is about AI or ML, the editor will ask you to make the weights of a trained model available.}

@@

\newpage
\section*{Acknowledgements}
Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, inessential words, or effusive comments. A person can be thanked for assistance, not “excellent” assistance, or for comments, not “insightful” comments, for example. Acknowledgements can contain grant and contribution numbers.

\section*{Author Contributions}
Author Contributions: Authors are required to include a statement to specify the contributions of each co-author. The statement can be up to several sentences long, describing the tasks of individual authors referred to by their initials. See the authorship policy page for further explanation and examples.

\section*{Author Information}
\textbf{Grey's Note} \textit{Two things are required in this Author Information section: (1) A statement about competing interests. I have no advice about what constitutes a competing interest, you will have to read about it and make your own decision. (2) A clear statement about who to contact with question about the paper. An example is below.}
% For example:
% The authors declare no competing interests. Please contact either the first author (Grey Nearing; \href{mailto:nearing@google.com}{nearing@google.com}) or Avinatan Hassidim; \href{mailto:avinatan@google.com}{avinatan@google.com}) for correspondence and requests, including questions regarding reprints and permissions.

\newpage
\section*{Extended Data}
\begin{enumerate}
    \item Table: scores of all metrics - \textbf{should I add this??}
    \item Detailed architecture? - \textbf{should I add this??}
    \item Colorful PDF examples for NZ and Japan
    \item Marginal distributions for NZ and Japan
    \item Colored scatter magnitude vs index, NZ and Japan - \textcolor{red}{\textbf{Done}}
    \item $m_c(t)$ for all three regions -  - \textcolor{green}{kind of finished plot, not sure}
    \item Bar plots of conditioned $<LL>$
    \item Spatial $m_c(x, y)$ and low $m_c(x, y) \leq m_c^{train}$ bins
\end{enumerate}

\newpage
\begin{figure}
    \centering
    \includegraphics[width=1\textwidth]{figures/detailed_architecture.pdf}
    \caption{\textbf{MAGNET architecture.} Described in detail in the methods section.}
    \label{fig:architecture}
\end{figure}

\newpage
\begin{figure}[h!]
	\centering
        \includegraphics[width=1\textwidth]{figures/info_gain_scatter_NZ_and_Japan.pdf}
	\caption{
            \textbf{Information gain over time domain for New Zealand and Japan data sets.} This figure presents results for the data sets not displayed in Fig. \ref{fig:info_gain_over_time}a. \textbf{a} Entire test set of New Zealand and a selected time span (\textbf{b}) focusing on the 2016 Te Araroa earthquake and the 2016 Kaikōura earthquake. \textbf{c} The entire Japan data set and a focused time span \textbf{d} of the 2011 Tōhoku earthquake and some of its aftershocks.
         }
\label{fig:intro_fig}
\end{figure}

\newpage

@@

>>>>

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/combined_batplots.pdf}
    \caption{
        \textbf{Metric scores for the MAGNET model, and common benchmark models. a}, \textbf{b}, \textbf{c}, Minus mean information content of our MAGNET model (red) and other common benchmark magnitude predictors (see labels on bars in the figure). A lower score indicated a better preforming model. \textbf{d, e, f}, Receiver Operating Characteristic (ROC) and \textbf{g, h, i} the interpolated precision-recall (PR) curve for a binary classifier determining the existence of a large ($m>=4$) event. The performance of such a classifier can be quantified by the area under the curve (AUC), noted in each frame, color coded identically to the bar plots. For the AUC metrics, a higher score indicated a better preforming classifier.
        \neri{Organize titles and font sizes}
        }
        \label{fig:metrics}
\end{figure}

@@

\subsection{Measurement of temporal incompleteness}
Temporal incompleteness is measured at the times of events by fitting the completeness magnitude to a window of 300 events: 150 past events, 1 current event, 149 future events. The window is constructed by the same algorithm, used to construct the \textit{Last $n$ events benchmark}.

\section*{Data and code availability}
The datasets analysed during the current study are available in the relevant references mentioned throught the article.
The code for analysing the data is available in \textcolor{red}{\textbf{**link TBD**}}
% This is a required section. Guidelines for data availability: \url{https://www.nature.com/documents/nr-data-availability-statements-data-citations.pdf}.

@@

\newpage
\section*{Acknowledgements}
\yohai{Legit to use your comment to give you tasks? TODO by \textbf{Yohai}}
Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, inessential words, or effusive comments. A person can be thanked for assistance, not “excellent” assistance, or for comments, not “insightful” comments, for example. Acknowledgements can contain grant and contribution numbers.

\section*{Author Contributions}
N.B., O.Z., O.G., Y.B.S.
N.B., O.Z., O.G. and Y.B.S. designed the research. N.B. conducted the experiments. N.B., O.G. and Y.B.S. preformed the statistical analysis. N.B. and O.Z. wrote the code base. N.B. and Y.B.S. wrote the manuscript. All authors reviewed the manuscript.

\section*{Author Information}
\yohai{Legit to use your comment to give you tasks? TODO by \textbf{Yohai} x2}
\textbf{Grey's Note} \textit{Two things are required in this Author Information section: (1) A statement about competing interests. I have no advice about what constitutes a competing interest, you will have to read about it and make your own decision. (2) A clear statement about who to contact with question about the paper. An example is below.}
% For example:
% The authors declare no competing interests. Please contact either the first author (Grey Nearing; \href{mailto:nearing@google.com}{nearing@google.com}) or Avinatan Hassidim; \href{mailto:avinatan@google.com}{avinatan@google.com}) for correspondence and requests, including questions regarding reprints and permissions.

\newpage
\section*{Extended Data}
\begin{enumerate}
    \item Table: scores of all metrics - \textcolor{red}{\textbf{should I add this??}}
    \item Detailed architecture? - \textcolor{red}{\textbf{Make a nice version}}
    \item Colorful PDF examples for NZ and Japan
    \item Marginal distributions for NZ and Japan
    \item Colored scatter magnitude vs index, NZ and Japan - \textcolor{blue}{\textbf{Done}}
    \item $m_c(t)$ for all three regions -  - \textcolor{blue}{Done}
    \item Bar plots of conditioned $<LL>$
    \item Spatial $m_c(x, y)$ and low $m_c(x, y) \leq m_c^{train}$ bins
\end{enumerate}

@@

\newpage
\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/detailed_architecture.pdf}
    \caption{\neri{TODO: this needs to be nicer} \textbf{MAGNET architecture.} Described in detail in the methods section.}
    \label{fig:architecture}
\end{figure}

\newpage
\begin{figure}[h!]
	\centering
        \includegraphics[width=1\textwidth]{figures/info_gain_scatter_NZ_and_Japan.pdf}
	\caption{
            \textbf{Information gain over time domain for New Zealand and Japan data sets.} This figure presents results for the data sets not displayed in Fig. \ref{fig:info_gain_over_time}a. \textbf{a} Entire test set of New Zealand and a selected time span (\textbf{b}) focusing on the 2016 Te Araroa earthquake and the 2016 Kaikōura earthquake. \textbf{c} The entire Japan data set and a focused time span \textbf{d} of the 2011 Tōhoku earthquake and some of its aftershocks.
         }
\label{fig:nz_japan_info_gain}
\end{figure}

\newpage
\begin{figure}[h!]
    \centering
        \includegraphics[width=1\textwidth]{figures/temporal_incompleteness.pdf}
    \caption{
    \textbf{Temporal Incompleteness for test sets of all regions examined.} Calculated temporal incompleteness for test sets is presented in red, superimposed over the events above (black) and below (grey) the completeness magnitude of the train set, $m_c^{(train)}$. Events are plotted by the index of their appearance in the test set. Events of low magnitude, $m_c^{(train)}$ (grey), that are not included in the evaluation, are plotted at their interpolated location along the horizontal axis. Data for Japan (c) is displayed only near the 2011 Tōhoku earthquake to reduce data density and enhance clarity of the figure.
    }
    \label{fig:temp_incompleteness}
\end{figure}
